---
title: "Elemental and Presentation UI Layers"
author: "Ian Lyttle & Alex Shum"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design of Shinypods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r prelim, echo=FALSE}
library("shinypod")

cat_r_file <- function(file){
  cat("```R", readLines(file), "```", sep = "\n")  
}

cat_r_fn <- function(fn){
  cat("```R", capture.output(eval(fn)), "```", sep = "\n")  
}
```

In this vignette, we go into a little more detail of the structure of a shinypod.

## Essentials

There are three essential functions for the UI. Let's consider the dygraph shinypod:

- `dygraph_ui_input()` returns a named `shiny::tagList` of input elements
- `dygraph_ui_output()` returns a named `shiny::tagList` of output elements
- `dygraph_ui_misc()` returns a named `shiny::tagList` of miscellaneous elements, like help guides.

### Convention

The functions `dygraph_sidebar_side()` and `dygraph_sidebar_main()` draw upon the elemental ui functions: `dygraph_ui_input()`, `dygraph_ui_output()` and `dygraph_ui_misc()`.

Let's look at the code for `dygraph_ui_input()`:

```{r dygraph_ui_input, comment="", echo=FALSE, results = "asis"}
cat_r_fn(dygraph_ui_input)
```

We see that this returns a named `tagList` of HTML elements.

Similarly for `dygraph_ui_output()`:

```{r dygraph_ui_output, comment="", echo=FALSE, results = "asis"}
cat_r_fn(dygraph_ui_output)
```

This returns an empty `tagList`, but we keep the function for completeness.

Finally for `dygraph_ui_misc()`:

```{r dygraph_ui_misc, comment="", echo=FALSE, results = "asis"}
cat_r_fn(dygraph_ui_misc)
```

## Presentation layer

On top of the elemental layer, we make a presentation layer that consists of the functions:

- `dygraph_sidebar_side()`
- `dygraph_sidebar_main()`

Here's `dygraph_sidebar_side()`:

```{r dygraph_sidebar_side, comment="", echo=FALSE, results = "asis"}
cat_r_fn(dygraph_sidebar_side)
```

Here's `dygraph_sidebar_main()`:

```{r dygraph_sidebar_main, comment="", echo=FALSE, results = "asis"}
cat_r_fn(dygraph_sidebar_main)
```

These are the functions that you would use - in putting togther a shiny app, I would suggest using these presentation functions.
