---
title: "Using Shinypods"
author: "Ian Lyttle & Alex Shum"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design of Shinypods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Shiny modules can make it much easier to design and build more-complex shiny apps. 

Imagine that you are designing an app that requires the user to upload a csv file to be parsed into a data frame. You will wish to use the data frame elsewhere in the app. Instead of writing the parser  from scratch, you can use a set of **shinypod** functions in your app.

Shinypod is simply an implementaton of a set of design guidelines to allow you to use, remix, and even build your own shiny modules. An advantage of using such guidelings is that we can all use each others' shinypods more quickly and reliably.

Before getting started, you are referred to RStudio's [article on shiny modules](http://shiny.rstudio.com/articles/modules.html).

## First example - CSV parser

This package, shinypod, offers functions that you can use to put a CSV parser into your app. 

Let's look at a `ui.R` file:

```{r parse_ui, eval=FALSE}
library("shiny")
library("shinyjs")
library("shinyBS")
library("shinypod")

shinyUI(
  fluidPage(
    useShinyjs(),
    titlePanel("CSV parser"),
    sidebarLayout(
      sidebarPanel(
        read_delim_sidebar_side("csv")
      ),
      mainPanel(
        read_delim_sidebar_main("csv")
      )
    )
  )
)
```

And a `server.R` file:

```{r parse_server, eval=FALSE}
library("shiny")
library("shinyjs")
library("shinyBS")
library("shinypod")

shinyServer(function(input, output, session) {

  rct_data <- callModule(module = read_delim_server, id = "csv")
})
```

For this implementation we call upon three functions.

- `foo_sidebar_side()`
- `foo_sidebar_main()`
- `foo_server()`

The first two functions each return a named `shiny::tagList` of UI elements that can be used in a sidebar layout. The side panel contains the inputs, and the main panel contains the outputs. 

The third function is used in the server function to return a reactive element. This reactive returns the parsed dataframe. 

To use all three, all you have to do is call each of the three functions using the same `id`, in this case `"csv"` - this is to keep the shiny namespace tidy.

You can see this app in action [here].

## Second example - CSV parser with dygraph

In this example, we will look at what you can do by combining shinypods. Here, we will combine the pod used to parse a csv with a pod used to build a dygraph.

Dygraphs are especially handy for visualizing time series. Often, time-series data may be available in a data frame (parsed from a csv), and we wish to visualize it. This is where we would use a set of dygraphs functions from shinypod.

We can build onto our previous example by adding a dyraph shinypod.

Here's a `ui.R` file:

```{r parse_dy_ui, eval=FALSE}
library("shiny")
library("shinyBS")
library("shinyjs")
library("shinypod")
library("dygraphs")
library("dplyr")

fluidPage(
  useShinyjs(),
  titlePanel("CSV parser and dygraph"),
  sidebarLayout(
    sidebarPanel(
      read_delim_sidebar_side("csv"),
      tags$hr(),
      dygraph_sidebar_side("dyg")
    ),
    mainPanel(
      read_delim_sidebar_main("csv"),
      dygraphOutput("csv_dyg")
    )
  )
)
```

Here's a `server.R` file:

```{r parse_ui_server, eval=FALSE}
library("shiny")
library("shinyBS")
library("shinyjs")
library("shinypod")
library("dygraphs")
library("dplyr")

function(input, output, session) {

  rct_data <- callModule(read_delim_server, "csv")

  rct_dyg <- callModule(dygraph_server, "dyg", data = rct_data)

  output$csv_dyg <- renderDygraph({
    rct_dyg() %>%
      dyOptions(useDataTimezone = TRUE)
  })

}
```

A few things to note:

- we use a different namespace id for the read_delim pod (`"csv"`) as for the dygraph pod (`"dyg"`).
- we supply the reactive returned by the `read_delim_server()` function as an argument to the `dygraph_server()` function. This is how the data is moved around.
- the `dygraph_server()` function returns a reactive that returns a dygraph. We use this to build the output for the app. Note also that the dygraph shinypod returns a basic dygraph; we can use functions from the dygraphs package to customize it.

Here's a [link] to a working example.





